Torrential Twitter? Measuring the Severity of Harassment when Canadian Female Politicians Tweet about Climate Change

### 2024-01-25 Outputs/Paper/Publication_version/paper_for_pub.qmd
start editing for publication

---
title: "Torrential Twitter: Climate Change, Female Politicians, and Harassment"
author: "Inessa De Angelis^[Faculty of Information, University of Toronto, Inessa.DeAngelis@mail.utoronto.ca]"
thanks: "Code and select data available at: https://github.com/InessaDeAngelis/torrential_twitter_bell_conf "
date: "today"
date-format: "long"
abstract: "Women in politics face considerable harassment. Known as “gendertrolling,” this harassment is part of a global trend of violence against women in politics. There are many significant previously documented cases of gendertrolling on public social media platforms such as Twitter. Framed through the theory of deliberative democracy, this study conducts quantitative content analysis on the replies received by three female politicians in Canada when Tweeting about climate change and the environment. These politicians are Deputy Prime Minister Chrystia Freeland (from the Liberal Party), Elizabeth May (co-Leader of the Green Party), and Laurel Collins (from the New Democratic Party). I then built a Bayesian hierarchical model to investigate the relationship between the severity of harassment and the type of account that a user has. The majority of replies contain some form of harassment, often questioning authority or name-calling/gender insults, and come from users that have spam or anonymous accounts. *This paper contains language and themes that some readers may find offensive.*"
format: pdf
toc: true
number-sections: true
bibliography: conference_references.bib
csl: apa.csl
---

@@

<!-- The past few decades have led to structural transformation and fragmentation of the idea of the public sphere, moving from having a singular, "official" public sphere to a multitude of smaller parallel ones, called "subaltern counterpublics" and networked publics mediated by technological affordances [@brunshighfield; @fraser1991; @boyd2010]. Technology reorganized the flow of information in society, distinguishing networked publics from subaltern counterpublics and the "official" public. Discourse traditionally happened in mediated spaces, but the rise of social media allows it to happen everywhere and is theoretically breaking down barriers of who has access to the walled garden of democratic discourse [@mediatingmisogyny; @trifiro]. Despite the potential formation of new publics online, the public which continues to deny the existence of climate change are largely white, male, and Christian and have free time (often dictated by patriarchal norms) to troll female politicians on Twitter whose actions as climate leaders do not fit traditional gender roles and conceptions of power [@Anshelm; @mccrightdunlap; @degeus; @mantilla2013]. Networked publics can be potent for democracy, undermining the authority of female politicians and risking the civility of democratic discourse. Consequently, women have and continue to be on the margins of the public sphere, both in physical spaces like the House of Commons and online, as underscored by an @ipu global survey which found that 82% of female politicians reported being subjected to psychological violence (including harassment, threats, and sexist comments). In Canada, specific examples of on-going marginalization can be seen through the lived realities of women in politics, with only one female Prime Minister, Rt. Hon. Kim Campbell in 1993, and in 2024, only 30% of MPs are female [@degeus; @hoc].  -->

@@

<!-- Diverging scholarly viewpoints exist when operationalizing and applying the theory of deliberative democracy. Some scholars view the theory in rigid terms, evaluating all characteristics and scenarios, such as political parties, systems, and institutions on a "yes" or "no" basis [@sokolon]. Sometimes, other scholars overlook or discount the role of social media and emerging technologies in creating networked publics which can drastically shape the quality and quantity of online democratic discourse. Moreover, other scholars view the theory of deliberative democracy on a scale - looking at the extent in which healthy democratic speech exists in specific contexts [@rossini]. Communication and media studies scholars are concerned about the nuances of discourse and the specific contexts in which it is situated and how algorithms and platform affordances are building and shaping these contexts [@bruns2023]. -->

@@

After manually coding the 4,139 Tweets using my Google Form and Codebook (see [Appendix]), I proceeded to the second round of cleaning the Tweets using R [@r]. I eliminated columns of data which were useful for determining the type of account but not relevant for the next steps of my analysis. For my severity of harassment data, I updated value 1 to "Positive," 2 to "Neutral," 3 to "Questioning authority," 4 to "Name-calling/Gender insults," 5 to "Vicious language," 6 to "Credible threats," and 7 to "Hate speech" based on my Codebook. Furthermore, for my type of account data, I changed value 1 to "Personal," 2 to "Professional," 3 to "Bots," 4 to "Spammers," 5 to "Anonymous," and 6 to "Suspended/deleted" based on my Codebook. I tested the cleaned data using `validate` [@validate], primarily validating variable types and ensuring the number of rows in the dataset matches the number of Tweets I coded. I then analyzed and visualized my data in R [@r], using functions from `knitr` [@rknitr], `kableExtra` [@rKableExtra], `Formattable` [@rFormattable], and `ggplot2` [@rGgplot2].

@@

# Acknowledgments {.unnumbered}
I am grateful for Rohan Alexander's helpful suggestions and feedback on drafts of this paper. This research is supported in part by funding from the Social Sciences and Humanities Research Council of Canada (Canada Graduate Scholarship - Master's) and Government of Ontario (Ontario Graduate Scholarship).

@@

  - My coding form is available [here](https://forms.gle/Limy51uAcvM3B7K66)

>>>>

---
title: "Torrential Twitter? Measuring the Severity of Harassment when Canadian Female Politicians Tweet about Climate Change"
author: "Inessa De Angelis^[Faculty of Information, University of Toronto, Inessa.DeAngelis@mail.utoronto.ca]"
thanks: "Code and select data available at: https://github.com/InessaDeAngelis/torrential_twitter_bell_conf "
date: "today"
date-format: "long"
abstract: "Research has shown that women in politics face considerable harassment. Known as “gendertrolling,” this harassment is part of a global trend of violence against women in politics. This study investigates the severity of harassment three Canadian female politicians receive when Tweeting about climate change and the types of accounts behind the replies. Results from the quantitative content analysis reveal that 86% of Tweets contained some form of harassment. I then built a Bayesian hierarchical model to investigate the relationship between the severity of harassment and the type of account that a user has. The results from my model suggest that despite differences in political affiliation and status, the three politicians are equally impacted by harassment when Tweeting about climate change. *This paper contains language and themes that some readers may find offensive.*"
format: pdf
toc: true
number-sections: true
bibliography: publication_references.bib
csl: apa.csl
---

@@

@@

@@

After manually coding the 4,139 Tweets using my Codebook (see [Appendix]), I proceeded to the second round of cleaning. I eliminated columns of data which were useful for determining the type of account but not relevant for the next steps of my analysis. For my severity of harassment data, I updated value 1 to "Positive," 2 to "Neutral," 3 to "Questioning authority," 4 to "Name-calling/Gender insults," 5 to "Vicious language," 6 to "Credible threats," and 7 to "Hate speech" based on my Codebook. Furthermore, for my type of account data, I changed value 1 to "Personal," 2 to "Professional," 3 to "Bots," 4 to "Spammers," 5 to "Anonymous," and 6 to "Suspended/deleted" based on my Codebook. I tested the cleaned data using `validate` [@validate], primarily validating variable types and ensuring the number of rows in the dataset matches the number of Tweets I coded. I then analyzed and visualized my data using functions from `knitr` [@rknitr], `kableExtra` [@rKableExtra], `Formattable` [@rFormattable], and `ggplot2` [@rGgplot2].

@@

# Declaration of Conflicting Interests {.unnumbered}
The author declares no potential conflicts of interest with respect to the research, authorship, and/or publication of this article.

# Acknowledgments {.unnumbered}
I am grateful for Rohan Alexander's helpful suggestions and feedback on drafts of this paper and for Elizabeth Dubois' guidance during the early conception of this study. Thank you to Sarah Xie and Allyson Cui for their research assistance.

# Funding {.unnumbered}
The author discloses receipt of the following financial support for the research, authorship, and/or publication of this article: This study was funded by the Social Sciences and Humanities Research Council of Canada, Canada Graduate Scholarship (id: 766-2022-0719) and the Government of Ontario, *2023-2024 Ontario Graduate Scholarship*.

@@
